On exponential sums arising from the classical doubling method

### 2025-02-07 paper.tex
Some corrections

\newcommand{\parabolicSection}{\Phi^{\left(z_1, \dots, z_c\right)}}
\newcommand{\intertwiningOperator}{M^{\left(z_1, \dots, z_c\right)}}
\newcommand{\holomorphicRepresentation}{\depthZeroRepresentation^{\left(z_1, \dots, z_c\right)}}
\newcommand{\WhittakerFunctional}[1]{\ell_{#1, \fieldCharacter}}
\newcommand{\gWhittakerFunctional}[2]{\ell_{#1, #2}}
\newcommand{\gSpehWhittakerFunctional}[3]{\ell_{\SpehRepresentation{#1}{#3}, \fieldCharacter_{\UnipotentRadicalForWss{#2}{#3}}}}
\newcommand{\gShortSpehWhittakerFunctional}[3]{\ell_{\SpehRepresentation{#1}{#3}}}
\newcommand{\GaussSum}[2]{\mathcal{G}\left(#1, #2\right)}
\newcommand{\dblGaussSum}[2]{\mathcal{G}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblGaussSumScalar}[2]{\mathrm{G}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualGaussSumScalar}[2]{\mathrm{g}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GKGaussSum}[3]{\mathcal{G}\left(#1 \times #2, #3\right)}
\newcommand{\GKGaussSumScalar}[3]{\mathrm{G}\left(#1 \times #2, #3\right)}
\newcommand{\fieldCharacterkc}[2]{\fieldCharacter_{\left({#2}^{#1}\right)}}
\newcommand{\ExoticKloosterman}{\mathrm{Kl}}
\newcommand{\GaussSumCharacter}[4]{\tau_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\IrrCusp}{\Irr_{\mathrm{cusp}}}
\newcommand{\convolutionWithCompactSupport}{\boldsymbol{\mathrm{R}}}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\artinScrier}{\operatorname{AS}}
\newcommand{\KloostermanSumClassFunction}{\mathcal{K}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}

@@

\subsubsection{General linear groups computation}
Suppose that $F \colon \GL_k\left(\finiteField\right) \to \cComplex$ is a class function. If $\chi \colon \multiplicativegroup{\finiteField} \to \multiplicativegroup{\cComplex}$ is a non-trivial character, we define $$\dblVirtualGaussSumScalar{F}{\chi} = q^{-\frac{k^2}{2}} \sum_{g \in \GL_k\left(\finiteField\right)} \Phi_{\chi}\left(g\right) F\left(g\right).$$
It is clear that for an irreducible representation $\tau$ of $\GL_k\left(\finiteField\right)$, we have $$\dblVirtualGaussSumScalar{\trace \tau}{\chi} = \dblGaussSumScalar{\tau}{\chi} \cdot \dim \tau.$$
Since the assignment $g \mapsto \Phi_{\chi}\left(g\right)$ depends only on the semisimple part of $g$, by \cite[Theorem in Section 1.2]{SaitoShinoda2000} we have that if $T \cong \multiplicativegroup{\finiteFieldExtension{k}}$ and if $\theta \colon \multiplicativegroup{\finiteFieldExtension{k}} \to \multiplicativegroup{\cComplex}$ is a character then
$$\dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} = \frac{\grpIndex{\GL_k\left(\finiteField\right)}{T}}{q^{\frac{k^2}{2}}} \sum_{\substack{t \in \multiplicativegroup{\finiteFieldExtension{k}}\\
t \ne 1}} \theta\left(t\right) \chi\left(\FieldNorm{k}{1}\left(t-1\right)\right).$$
Using \Cref{thm:gln-doubling-gauss-sum-in-terms-of-kondo} and the fact that Kondo's Gauss sum is the same for every subrepresentation of $R_{T, \theta}$, we have the equality $$\dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} = \left(-1\right)^k \theta\left(-1\right) q^{-k} R_{T, \theta}\left(1\right) \tau\left(\chi^2, \fieldCharacter\right)^k \tau\left(\theta, \fieldCharacter\right) \tau\left(\theta^{-1} \times \chi^{-1}, \fieldCharacter\right).$$

@@

Hence we get
$$\sum_{\substack{x \in \multiplicativegroup{\finiteFieldExtension{k}}\\
		x \ne 1}} \theta \left(x\right) \chi\left(\FieldNorm{k}{1}\left(x - 1\right)\right) \chi\left(\FieldNorm{k}{1}\left(x^{-1} - 1\right)\right) = \frac{q^{\frac{k^2}{2}}}{\grpIndex{\GL_k\left(\finiteField\right)}{T}} \dblVirtualGaussSumScalar{R_{T, \theta_{\chi^{-1}}}}{\chi^2},$$
which by the results of the previous section equals
%$$R_{T,\theta}\left(1\right) = \left(-1\right)^{k-1} \frac{\grpIndex{\GL_k\left(\finiteField\right)}{T}}{q^{\frac{k^2 - k}{2}}}$$
$$-\theta\left(-1\right) q^{-\frac{k}{2}} \tau\left(\chi^4, \fieldCharacter\right)^k \tau\left(\theta \times \chi^{-1}, \fieldCharacter\right) \tau\left(\theta^{-1} \times \chi^{-1}, \fieldCharacter\right).$$

@@

If $\alpha \colon \multiplicativegroup{\finiteFieldExtension{n}} \to \multiplicativegroup{\cComplex}$ is a character, denote $$\tau\left(\alpha, \fieldCharacter_n\right) = -q^{-\frac{n}{2}}\sum_{x \in \multiplicativegroup{\finiteFieldExtension{n}}} \alpha\left(x\right) \fieldCharacter_n\left(x\right),$$
where $\fieldCharacter_n \colon \finiteFieldExtension{n} \to \multiplicativegroup{\cComplex}$ is given by $\fieldCharacter_n = \fieldCharacter \circ \trace_{\finiteFieldExtension{n} : \finiteField}$. If $\chi \colon \multiplicativegroup{\finiteField} \to \multiplicativegroup{\cComplex}$ is a character and if $\alpha$ is as above, denote
$$\tau\left(\alpha \times \chi, \fieldCharacter_n\right) = -q^{-\frac{n}{2}}\sum_{x \in \multiplicativegroup{\finiteFieldExtension{n}}} \alpha\left(x\right) \chi\left( \FieldNorm{n}{1}\left(x\right)\right) \fieldCharacter_n\left(x\right),$$

@@

Let $T \cong \finiteFieldExtension{2n_1}^1 \times \finiteFieldExtension{2n_2}^1 \times \dots \times \finiteFieldExtension{2n_r}^1 \times \multiplicativegroup{\finiteFieldExtension{m_1}} \times \dots \multiplicativegroup{\finiteFieldExtension{m_s}}$ be a torus such that $\sum_{i=1}^r 2n_i + \sum_{j=1}^s 2m_j = \dim_{\finiteField} \hermitianSpace$. Let $\theta = \alpha_1 \times \dots \times \alpha_r \times \beta_1 \times \dots \times \beta_s$. Consider the Deligne--Lusztig induction $\DeligneLusztigInduction{T}{G}\theta$. Write $$\DeligneLusztigInduction{T}{G}\theta = \sum_{\pi} c_{\pi} \cdot \trace \pi,$$
where $\pi$ goes over all the irreducible representations of $G$ and $c_{\pi} \in \zIntegers$. Then for any irreducible $\pi$ such that $c_{\pi} \ne 0$, and any $\chi$ such that $\chi^2 \ne 1$, we have
$$\dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = q^{-\frac{\dim \hermitianSpace}{2}} \prod_{j=1}^r \GaussSumScalar{\alpha'_j \times \chi^{-1}}{\fieldCharacter_{2n_j}} \prod_{i=1}^s \GaussSumScalar{\beta_i \times \chi^{-1}}{\fieldCharacter_{m_i}} \GaussSumScalar{\beta_i^{-1} \times \chi^{-1}}{\fieldCharacter_{m_i}}.$$

>>>>

\newcommand{\parabolicSection}{\Phi^{\left(z_1, \dots, z_c\right)}}
\newcommand{\intertwiningOperator}{M^{\left(z_1, \dots, z_c\right)}}
\newcommand{\holomorphicRepresentation}{\depthZeroRepresentation^{\left(z_1, \dots, z_c\right)}}
\newcommand{\WhittakerFunctional}[1]{\ell_{#1, \fieldCharacter}}
\newcommand{\gWhittakerFunctional}[2]{\ell_{#1, #2}}
\newcommand{\gSpehWhittakerFunctional}[3]{\ell_{\SpehRepresentation{#1}{#3}, \fieldCharacter_{\UnipotentRadicalForWss{#2}{#3}}}}
\newcommand{\gShortSpehWhittakerFunctional}[3]{\ell_{\SpehRepresentation{#1}{#3}}}
\newcommand{\GaussSum}[2]{\mathcal{G}\left(#1, #2\right)}
\newcommand{\dblGaussSum}[2]{\mathcal{G}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\GaussSumScalar}[2]{\mathrm{G}\left(#1, #2\right)}
\newcommand{\dblGaussSumScalar}[2]{\mathrm{G}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblVirtualGaussSumScalar}[2]{\mathrm{g}^{\mathrm{dbl}}\left(#1, #2\right)}
\newcommand{\dblGammaFactor}[3]{\Gamma^{\mathrm{dbl}}\left(#1 \times #2, #3\right)}
\newcommand{\dblGammaFactorSpace}[4]{\Gamma^{\mathrm{dbl}}_{#1}\left(#2 \times #3, #4\right)}
\newcommand{\GKGaussSum}[3]{\mathcal{G}\left(#1 \times #2, #3\right)}
\newcommand{\GKGaussSumScalar}[3]{\mathrm{G}\left(#1 \times #2, #3\right)}
\newcommand{\fieldCharacterkc}[2]{\fieldCharacter_{\left({#2}^{#1}\right)}}
\newcommand{\ExoticKloosterman}{\mathrm{Kl}}
\newcommand{\GaussSumCharacter}[3]{\tau\left(#1 \times #2, #3\right)}
\newcommand{\IrrCusp}{\Irr_{\mathrm{cusp}}}
\newcommand{\convolutionWithCompactSupport}{\boldsymbol{\mathrm{R}}}
\newcommand{\ladicnumbers}{\algebraicClosure{\mathbb{Q}_{\ell}}}
\newcommand{\artinScrier}{\operatorname{AS}}
\newcommand{\KloostermanSumClassFunction}{\mathcal{K}}
\newcommand{\IsometryGroup}{\mathrm{Isom}}
\newcommand{\lieAlgebra}{\mathfrak{g}}
\newcommand{\DeligneLusztigInduction}[2]{\mathrm{R}_{#1}^{#2}}
\newcommand{\algebraicGroup}[1]{\boldsymbol{\mathrm{#1}}}

@@

\subsubsection{General linear groups computation}
Suppose that $F \colon \GL_k\left(\finiteField\right) \to \cComplex$ is a class function. If $\chi \colon \multiplicativegroup{\finiteField} \to \multiplicativegroup{\cComplex}$ is a non-trivial character, we define $$\dblVirtualGaussSumScalar{F}{\chi} = q^{-\frac{k^2}{2}} \sum_{g \in \GL_k\left(\finiteField\right)} \Phi_{\chi}\left(g\right) F\left(g\right).$$
It is clear that for an irreducible representation $\tau$ of $\GL_k\left(\finiteField\right)$, we have $$\dblVirtualGaussSumScalar{\trace \tau}{\chi} = \dblGaussSumScalar{\tau}{\chi} \cdot \dim \tau.$$
Since the assignment $g \mapsto \Phi_{\chi}\left(g\right)$ depends only on the semisimple part of $g$, by \cite[Theorem in Section 1.2]{SaitoShinoda2000} we have that if $T \cong \multiplicativegroup{\finiteFieldExtension{k}}$ and if $\theta \colon \multiplicativegroup{\finiteFieldExtension{k}} \to \multiplicativegroup{\cComplex}$ is a character then
$$\dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} = \frac{\grpIndex{\GL_k\left(\finiteField\right)}{T}}{q^{\frac{k^2}{2}}} \sum_{\substack{t \in \multiplicativegroup{\finiteFieldExtension{k}}\\
t \ne 1}} \theta\left(t\right) \chi\left(\FieldNorm{k}{1}\left(t-1\right)\right).$$
Using \Cref{thm:gln-doubling-gauss-sum-in-terms-of-kondo} and the fact that Kondo's Gauss sum is the same for every subrepresentation of $R_{T, \theta}$, we have the equality $$\dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} = \left(-1\right)^k \theta\left(-1\right) q^{-\frac{3k}{2}} R_{T, \theta}\left(1\right) \tau\left(\chi, \fieldCharacter\right)^k \tau\left(\theta, \fieldCharacter\right) \tau\left(\theta^{-1} \times \chi^{-1}, \fieldCharacter\right).$$

@@

Hence we get
$$\sum_{\substack{x \in \multiplicativegroup{\finiteFieldExtension{k}}\\
		x \ne 1}} \theta \left(x\right) \chi\left(\FieldNorm{k}{1}\left(x - 1\right)\right) \chi\left(\FieldNorm{k}{1}\left(x^{-1} - 1\right)\right) = \frac{q^{\frac{k^2}{2}}}{\grpIndex{\GL_k\left(\finiteField\right)}{T}} \dblVirtualGaussSumScalar{R_{T, \theta_{\chi^{-1}}}}{\chi^2},$$
which by the results of the previous section equals
%$$\frac{q^{\frac{k^2}{2}}}{\grpIndex{\GL_k\left(\finiteField\right)}{T}} \left(-1\right)^k \theta\left(-1\right) q^{-\frac{3k}{2}} R_{T, \theta}\left(1\right) \tau\left(\chi^2, \fieldCharacter\right)^k \tau\left(\theta, \fieldCharacter\right) \tau\left(\theta^{-1} \times \chi^{-1}, \fieldCharacter\right). $$
%$$\frac{q^{\frac{k^2}{2}}}{\grpIndex{\GL_k\left(\finiteField\right)}{T}} \left(-1\right)^k \theta\left(-1\right) q^{-\frac{3k}{2}} \left(-1\right)^{k-1} \frac{\grpIndex{\GL_k\left(\finiteField\right)}{T}}{q^{\frac{k^2 - k}{2}}} \tau\left(\chi^2, \fieldCharacter\right)^k \tau\left(\theta, \fieldCharacter\right) \tau\left(\theta^{-1} \times \chi^{-1}, \fieldCharacter\right).$$
%$$R_{T,\theta}\left(1\right) = \left(-1\right)^{k-1} \frac{\grpIndex{\GL_k\left(\finiteField\right)}{T}}{q^{\frac{k^2 - k}{2}}}$$
$$-\theta\left(-1\right) q^{-k} \tau\left(\chi^2, \fieldCharacter\right)^k \tau\left(\theta \times \chi^{-1}, \fieldCharacter\right) \tau\left(\theta^{-1} \times \chi^{-1}, \fieldCharacter\right).$$

@@

If $\alpha \colon \multiplicativegroup{\finiteFieldExtension{n}} \to \multiplicativegroup{\cComplex}$ is a character, denote $$\tau\left(\alpha, \fieldCharacter_n\right) = -q^{-\frac{n}{2}}\sum_{x \in \multiplicativegroup{\finiteFieldExtension{n}}} \alpha\left(x\right) \fieldCharacter_n\left(x\right),$$
where $\fieldCharacter_n \colon \finiteFieldExtension{n} \to \multiplicativegroup{\cComplex}$ is given by $\fieldCharacter_n = \fieldCharacter \circ \trace_{\finiteFieldExtension{n} \slash \finiteField}$. If $\chi \colon \multiplicativegroup{\finiteField} \to \multiplicativegroup{\cComplex}$ is a character and if $\alpha$ is as above, denote
$$\tau\left(\alpha \times \chi, \fieldCharacter_n\right) = -q^{-\frac{n}{2}}\sum_{x \in \multiplicativegroup{\finiteFieldExtension{n}}} \alpha\left(x\right) \chi\left( \FieldNorm{n}{1}\left(x\right)\right) \fieldCharacter_n\left(x\right),$$

@@

Let $T \cong \finiteFieldExtension{2n_1}^1 \times \finiteFieldExtension{2n_2}^1 \times \dots \times \finiteFieldExtension{2n_r}^1 \times \multiplicativegroup{\finiteFieldExtension{m_1}} \times \dots \multiplicativegroup{\finiteFieldExtension{m_s}}$ be a torus such that $\sum_{i=1}^r 2n_i + \sum_{j=1}^s 2m_j = \dim_{\finiteField} \hermitianSpace$. Let $\theta = \alpha_1 \times \dots \times \alpha_r \times \beta_1 \times \dots \times \beta_s$. Consider the Deligne--Lusztig induction $\DeligneLusztigInduction{T}{G}\theta$. Write $$\DeligneLusztigInduction{T}{G}\theta = \sum_{\pi} c_{\pi} \cdot \trace \pi,$$
where $\pi$ goes over all the irreducible representations of $G$ and $c_{\pi} \in \zIntegers$. Then for any irreducible $\pi$ such that $c_{\pi} \ne 0$, and any $\chi$ such that $\chi^2 \ne 1$, we have
$$\dblGammaFactorSpace{\hermitianSpace}{\pi}{\chi}{\fieldCharacter} = q^{-\frac{\dim \hermitianSpace}{2}} \prod_{j=1}^r \GaussSumCharacter{\alpha'_j}{\chi^{-1}}{\fieldCharacter_{2n_j}} \prod_{i=1}^s \GaussSumCharacter{\beta_i}{\chi^{-1}}{\fieldCharacter_{m_i}} \GaussSumCharacter{\beta_i^{-1}}{\chi^{-1}}{\fieldCharacter_{m_i}}.$$
